---
title: "Ploty"
output: 
  html_document:
    code_folding: hide
    toc: true
    toc_float: true
---

```{r message = FALSE, echo = FALSE}
library(tidyverse)
library(plotly)
```

*CREATING THE PLOTLY WITH THE EXPLORATORY ANALYSIS FILE, NOT FINAL, STILL NEED TO ADD AND EDIT* 


**General Trend for each PUMA in each borough: Vaccination Rate, Death Rate, Hospitalization Rate for each PUMA in each Borough**

# Hospitalization rates across PUMAs, colored by borough
```{r eval = FALSE}
puma_hosp_highest10 <-
  puma_level_data %>% 
  mutate(puma = fct_reorder(puma, covid_hosp_rate)) %>% 
  mutate(text_label = str_c("PUMA: ", puma, 
                            "\n% Hospitalization: ", covid_hosp_rate)) %>%
  plot_ly(y = ~covid_hosp_rate, 
          x = ~puma, 
          color = ~borough, 
          type = "scatter",
          mode = "markers", 
          colors = "viridis",
          text = ~ text_label) %>% 
  config(scrollZoom = FALSE,
         modeBarButtonsToAdd = list('drawline', 
                                 'drawopenpath', 
                                 'drawclosedpath', 
                                 'drawcircle', 
                                 'drawrect', 
                                 'eraseshape'),
         displayModeBar = TRUE) %>%
  layout(plot_bgcolor = '#e5ecf6',
          title = "Hospitalization rates across PUMAs", 
          xaxis = list(
            title = "PUMA", 
            tickangle = 60,
            zerolinecolor = '#ffff',
            zerolinewidth = 2,
            gridcolor = 'ffff'),
          yaxis = list(
            title = "% Hospitalized", 
            zerolinecolor = '#ffff',
            zerolinewidth = 2,
            gridcolor = 'ffff'),
         legend = list(title = list(text = 'Borough')))
```


# Death rates across PUMAs, colored by borough
```{r include = FALSE}

```


# Vax rates across PUMAs, colored by borough
```{r include = FALSE}

```

**Which PUMAs have the highest Hospitalization Rate, Death Rate, Vaccination Rate**

# Puma with the highest hospitalization rate 
```{r include = FALSE}

```


# Puma with the highest Death rate 
```{r include = FALSE}

```


# Puma with the highest Vaccination rate 
```{r include = FALSE}

```


**Which PUMAs has the lowest Hospitalization Rate, Death Rate, Vaccination Rate**


# PUMAs with lowest hospitalization rates
```{r include = FALSE}

```


# PUMAs with lowest death rates
```{r include = FALSE}

```



# PUMAs with the lowest Vaccination rates
```{r include = FALSE}

```

