---
title: "Ploty"
output: 
  html_document:
    code_folding: hide
    toc: true
    toc_float: true
---

```{r message = FALSE, echo = FALSE}
library(tidyverse)
library(plotly)
```

*CREATING THE PLOTLY WITH THE EXPLORATORY ANALYSIS FILE, NOT FINAL, STILL NEED TO ADD AND EDIT* 


**General Trend for each PUMA in each borough: Vaccination Rate, Death Rate, Hospitalization Rate for each PUMA in each Borough**

# Hospitalization rates across PUMAs, colored by borough
```{r include = FALSE}

```


# Death rates across PUMAs, colored by borough
```{r include = FALSE}

```


# Vax rates across PUMAs, colored by borough
```{r include = FALSE}

```

**Which PUMAs have the highest Hospitalization Rate, Death Rate, Vaccination Rate**

# Puma with the highest hospitalization rate 
```{r include = FALSE}

```


# Puma with the highest Death rate 
```{r include = FALSE}

```


# Puma with the highest Vaccination rate 
```{r include = FALSE}

```


**Which PUMAs has the lowest Hospitalization Rate, Death Rate, Vaccination Rate**


# PUMAs with lowest hospitalization rates
```{r include = FALSE}

```


# PUMAs with lowest death rates
```{r include = FALSE}

```



# PUMAs with the lowest Vaccination rates
```{r include = FALSE}

```

